### 2021-11-24 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@2056db570ad5def0d6b14696f23b9cb0d263a24f 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Thesis proposal}
\vskip 5em

\href{https://orcid.org/0000-0002-6506-6487}{Michael D.\,Catchen}%

\textsuperscript{1}\,McGill University\quad \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences

\textbf{Correspondance to:}\\
Michael D. Catchen --- \texttt{michael.catchen@mail.mcgill.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
The proposal for my thesis, \emph{Simulation models for predictive
ecology}

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Within the last several hundred years, human activity has induced rapid
changes in Earth's atmosphere, oceans, and surface. Greenhouse gas
emissions have caused an increase the temperature of both Earth's
terrain and oceans, and both agricultural and urban development has
rapidly reshaped Earth's land cover. The bulk of this change has
occurred within the last several hundred years, a geological instant,
inducing a sudden shift in conditions to Earth's climate and biosphere.
As a result \emph{ecological forecasting}---modeling how ecosystems and
their services will change in the future---and then using these
forecasts to make decisions to mitigate the negative consequences of
this change on ecosystems, their functioning, and the services they
provide to humans has emerged as an imperative for ecology and
environmental science (Dietze 2017). However, robust prediction of
ecological processes is, to say the least, quite difficult (Beckage
\emph{et al.} 2011; Petchey \emph{et al.} 2015). This difficultly is
compounded by a few factors, the first being that sampling ecosystems is
not easy. Ecological data is often biased, noisey, and sparse in both
space and time. The current paucity of ecological data has resulted in
much interest in developing global systems for \emph{ecosystem
monitoring} (Makiola \emph{et al.} 2020), which would systematize the
collection of biodiversity data in a manner that makes detecting and
predicting change more possible than at the moment (Urban \emph{et al.}
2021).

The second major challenge in ecological forecasting is that the
underlying dynamics of ecological processes are unknown and instead must
be inferred from this (sparse) data. Much of the history of
quantitatively modeling ecosystems have been done in the language of
dynamical systems, describing how the value of an observable state of
the system, represented by a vector of numbers
\([x_1, x_2, \dots, x_n]^T = \vec{x}\) changes over time, yielding
models in the form of differential equations in continuous-time
settings, \(\frac{dx}{dt} = f(x)\), or difference equations in
discrete-time settings, \(x_t = f(x_{t-1})\), where
\(f:\mathbb{R}^n \to \mathbb{R}^n\) is an arbitrary function describing
how the system changes on a moment-to-moment basis (e.g.~in the context
of communities, \(f\) could be Lotka-Voltera, Holling-Type-II or
DeAngelis-Beddington functional response). The form of this functional
response in real systems, and whether it is meaningfully non-zero for a
given species interaction, is effectively unknown and must be predicted
(Strydom \emph{et al.} 2021b), and some forms of these dynamics are
inherently more ``forecastable'' than others (Beckage \emph{et al.}
2011; Chen \emph{et al.} 2019; Pennekamp \emph{et al.} 2019). The
initial success of these forms of models can be traced back to the
larger program of ontological reductionism, which became the default
approach to modeling in the sciences after its early success in physics,
which, by the time ecology was becoming a quantitative science (sometime
in the 20th century, depending on who you ask), became the foundation
for mathematical models in ecology.

However, we run into many problems when aiming to apply this type of
model to empirical ecological data. Ecosystems are perhaps the
quintessential example of system that cannot be understood by iterative
reduction of its components into constituent parts---ecological
phenomena are emergent: the product of different mechanisms operating at
different spatial, temporal, and organizational scales (Levin 1992).
Further this analytical approach to modeling explicitly ignores known
realities: ecological dynamics not deterministic and many analytic
models in ecology assume long-run equilibrium. Finally, perhaps the
biggest challenge in using these models to describe ecological processes
is ecosystems consist of more dimensions than the tools of analytic
models are suited for. As the number of variables in a model increases,
so does the ability of the scientist to discern clear relationships
between them given a fixed amount of data, the so-called ``curse'' of
dimensionality.

But these problems are not solely unique to ecology. The term
\emph{ecological forecasting} implicitly creates an analogy with weather
forecasting. Although it has become a trite joke to complain about the
weather forecast being wrong, over the least 50 years the field of
numerical weather prediction (NWP) has dramatically improved our ability
to predict weather across the board (Bauer \emph{et al.} 2015). The
success of NWP, and the Earth observation systems that support it (Hill
\emph{et al.} 2004), should serve as a template for development of a
system for monitoring Earth's biodiversity. Much like ecology, NWP is
faced with high-dimensional systems that are governed by different
mechanisms at different scales. The success of NWP is that, rather than,
say, attempt to forecast the weather in Québec by applying Navier-Stokes
to entire province, to instead use simulation models which describe
known mechanisms at different scales, and use the availability to
increasing computational power to directly simulate many batches of
dynamics which directly incorporate stochasticity and uncertainty in
parameter estimates via random number generation.

But forecasting is only half the story---if indeed ``{[}ecologists{]}
have hitherto only interpreted the world in various ways; the point is
to change it,'' then once we have a forecast about how an ecosystem will
change in the future, what if this forecast predicts a critical
ecosystem service will deteriorate? We are still left with the question,
what do we in the time being to mitigate the potentially negative
consequences a forecast predicts? In this framing, mitigating the
consequences of anthropogenic change on ecosystems becomes an
optimization problem: given a forecast of the future state of the
system, and some ``goal'' state for the future, the problem is then to
optimize our intervention into the system to maximize the probability
the system approaches our ``goal'' state. This dissertation aims to
provide a framework for ecosystem monitoring and forecasting
(fig.~\ref{fig:thesis}, left), and each chapter address some aspect of
this pipeline to data from a monitoring network to forecasts to
mitigation strategy (fig.~\ref{fig:thesis}, right).

\begin{figure}
\hypertarget{fig:thesis}{%
\centering
\includegraphics{./figures/thesisconcept.png}
\caption{Left: a framework for ecosystem monitoring, forecasting, and
mitigation. From the top, a set of biodiversity observatories which form
a monitoring network. Each collect various biodiversity data products.
From this raw data, we derive essential biodiversity variables (EBVs),
and forecast how they change over time. Based on this forecast, we wish
to choose the best possible mitigation strategy to maximize the chance
the realized future outcomes of the EBV approaches the target state.
Right: Each panel represents a chapter of the thesis, which follows the
flow of the framework on the left.}\label{fig:thesis}
}
\end{figure}

The primary research challenges this thesis addresses are: how do we
design ecological samples to? How do we build the software
infrastructure to assimilate data from a variety of sources? And how do
we propagate uncertainty from data to forecasts? The flow of chapters
follows the flow in fig.~\ref{fig:thesis} (left), from data collection
via a monitoring network, to forecasting an essential biodiversity
variable (EBV), to optimizing mitigation strategy based on constraints.
In chapter one, we discuss how simulation can aid in the design of
ecological samples and monitoring network design. In chapter two we use
data to forecast the uncoupling of a plant-pollinator network. In
chapter three, we apply simulation methods in landscape ecology to
optimize corridor placement to maximize the time-until-extinction of a
metapopulation. The fourth and final chapter is the software
(\emph{MetacommunityDynamics.jl}) which enables the rest of the
dissertation.

\hypertarget{chapter-one-optimizing-spatial-sampling-of-species-interactions}{%
\section{Chapter One: Optimizing spatial sampling of species
interactions}\label{chapter-one-optimizing-spatial-sampling-of-species-interactions}}

\hypertarget{objective}{%
\subsection{Objective}\label{objective}}

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
observing an interaction between species in order to aid in the design
of samples of ecological interactions across space, and to provide a
null expectation of the false-negative probability for a dataset of a
given size. Further it then proposes a method for optimizing the spatial
sampling locations to maximize the probability of detecting an
interaction between two species given a fixed number of total of
observations and the distributions of each species. This addresses the
optimization of monitoring network part of the flow from data to
mitigation at the top of fig.~\ref{fig:thesis}, left. There is more than
one way to observe a false-negative when sampling interactions---
fig.~\ref{fig:fnrtaxonomy} shows a taxonomy of false-negatives in
occurrence, co-occurrence, and interaction data. The goal of this
chapter is to untangle the relationship between co-occurrence
false-negatives (and their relation to relative abundance) and
interaction false-negatives.

\begin{figure}
\hypertarget{fig:fnrtaxonomy}{%
\centering
\includegraphics{./figures/ch2.png}
\caption{A taxonomy of occurrence, co-occurrence, and interaction false
negatives in data}\label{fig:fnrtaxonomy}
}
\end{figure}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

We begin by proposing a method to compute a null expectation of the
probability of an interaction false-negative as a function number of
total observations of individuals of \emph{all species in the species
pool}. This is done by simulating the process of observation, where the
probability that an observation is of a given species is that species'
relative abundance. We show that the realized false-negative rate can be
quite high, simply as a byproduct of the distribution of relative
abundance in communities. We use a log-normal distribution of relative
abundance (Hubbell 2001) and simulating the process of observation on
food-webs generated using the niche model (Williams \& Martinez 2000)
with connectance parameterized by the flexible-links model (MacDonald
\emph{et al.} 2020). An example of this relation for networks with
varying species richness is shown in fig.~\ref{fig:fnr}.

\begin{figure}
\centering
\includegraphics{./figures/negativebinomial.png}
\caption{Relationship between total number of observed interactions
negatives (x-axis) and the probability that an interaction is a
true-negative as a function of different realized false-negative rates
(colors), assuming each observed negative is independent---the same as
the negative-bernoulli distribution.}
\end{figure}

We then go on to test some assumptions of this neutral model with
empirical data. Primarily we analytically show that our neutral model,
if anything, underestimates the probability of false-negatives if there
are positive associations between species co-occurrence, and we show
these positive associations exist in two sets of spatially replicated
samples of interaction networks (Thompson \& Townsend 2000; Hadfield
\emph{et al.} 2014), fig.~\ref{fig:posassoc}---further I'm planning to
add the field data from chapter two into this analysis once available.
Finally this chapter proposes a simulated annealing method to optimize
the a set of \(n\) points in space to maximize the probability of
detecting an interaction between two species \(a\) and \(b\) with
\emph{known} distributions \(D_a\), \(D_b\).

\hypertarget{results}{%
\subsection{Results}\label{results}}

The first major result is using the simulation of the observation
process described above to generate expectations of interaction
false-negative rate (FNR) as a function of total number of observations,
with the goal being for this estimate to be used as correction for
detection error when fitting an interaction prediction model. This
relationship varies with the total richness of the metaweb (see
fig.~\ref{fig:fnr}).

\begin{figure}
\hypertarget{fig:fnr}{%
\centering
\includegraphics{./figures/ch2_fnr.png}
\caption{The realized false-negative-rate of interaction detection
(y-axis) as a function of the total number of observations of all
species in the species pool (different richnesses in different colors).
Each point is the mean of 50 replicates, with one standard-deviation in
the first shade, and two standard deviations in the second
shade.}\label{fig:fnr}
}
\end{figure}

The second major result is that we analytically show that the this
simulated observation model, by assuming that there is no association
between observing two species given that they interact, actually under
predicts the realized false-negative interaction rate. We then
demonstrate that this positive association association exists in two
empirical systems fig.~\ref{fig:posassoc}.

\begin{figure}
\hypertarget{fig:posassoc}{%
\centering
\includegraphics{./figures/positiveassociations.png}
\caption{Demonstrates positive associations in co-occurrence. Left: the
product of the marginal probability of observing two species (A and B)
in a sample (x-axis) agaianst the computed joint probability of
observing these species together (y-axis). Dashed line indicates
\(y=x\), meaning no association between the two. Each point is an
observed interaction between two species. Right: the distribution of the
difference between these joint and marginal probabilities. Both are
non-zero with \(p < 10^-50\) via a t-test.}\label{fig:posassoc}
}
\end{figure}

\hypertarget{progress}{%
\subsection{Progress}\label{progress}}

This chapter is mostly complete. The only remaining work is the
implementation of simulated annealing optimization process. This will be
done by using a proposal function which takes a set of coordinates in
space and proposes a new location for each point based on a
distance-decaying kernel.

\hypertarget{chapter-two-forecasting-the-spatial-uncoupling-of-a-plant-pollinator-network}{%
\section{Chapter Two: Forecasting the spatial uncoupling of a
plant-pollinator
network}\label{chapter-two-forecasting-the-spatial-uncoupling-of-a-plant-pollinator-network}}

\hypertarget{objective-1}{%
\subsection{Objective}\label{objective-1}}

Interactions between plants and pollinators form networks which together
structure the ``architecture of biodiversity'' (Bascompte \& Jordano
2007). The functioning and stability of ecosystems emerge from these
interactions, but anthropogenic change threatens to unravel and
``rewire'' these interaction networks (CaraDonna \emph{et al.} 2017),
jeopardizing the persistence of these systems. Plant-pollinator networks
face two possible forms of rewiring in response to anthropogenic
environmental change: spatial and temporal. Range shifts could cause
interacting species to no longer overlap in space, and shifts in
phenology could cause interacting species to no longer occur at the same
time of year. This chapter uses several years of data on
bumblebee-flower phenology and interactions across several field sites,
each consisting of several plots across an elevational gradient,
combined with spatial records of species occurrence via GBIF to forecast
the uncoupling of the plant-pollinator metaweb of Colorado.

\begin{figure}
\centering
\includegraphics{./figures/ch1.png}
\caption{Chapter One conceptual figure. Left: the sources of data and
how they can be combined. Right: The flow from data to interaction
prediction using a few different interaction prediction models.}
\end{figure}

\hypertarget{methods-1}{%
\subsection{Methods}\label{methods-1}}

The data for this chapter is derived from multiple sources that can be
split into four categories. (1) Field data from three different field
sites across Colorado, each with multiple plots across an elevational
gradient, for seven, seven, and three years respectively. This data was
collected by Paul CaraDonna and Jane Oglevie (from the Rocky Mountain
Biological Laboratory; RMBL) and Julian Resasco (CU Boulder). (2) GBIF
spatial occurrence records of each of these species across Colorado,
including a metaweb of interactions across all of Colorado taken from
GBIF. (3) Remotely sensed data consisting of current and forecasted
bioclimatic variables from CHELSA. (4) Phylogenies for both bee and
flower species derived from NCBI GenBank barcodes for mitochondrial COI
(bumblebees) and chloroplast rbcL (flowers).

As the data we have is spatially sparse and likely to contain many
interaction false-negatives, we begin by predicting a metaweb of
interactions across Colorado as they exist \emph{in the present}. We do
this using a set of candidate interaction prediction models: relative
abundance only, phylogenetic embedding only (a la Strydom \emph{et al.}
(2021a)), niche embedding only (encompassing both phenology and
environmental niche, similar to Gravel \emph{et al.} (2019)), and all
pairwise combinations of those constituent models. After validating and
selecting the best performing model, we then predict how these
distributions of each of these species will change under the CMIP6
consensus climate forecast (Karger \emph{et al.} 2017), and then finally
quantify the reduction in spatial overlap between species for which
there is a predicted interaction.

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

The in-progress results are the prerequisites for the analysis outlined
above: phylogenies for both plant and bee species (appendix figure one)
and species distribution models for all species (an example shown in
appendix figure two).

\hypertarget{progress-1}{%
\subsection{Progress}\label{progress-1}}

At the moment, we have derived phylogenies and SDMs for all the species
present in the Colorado GBIF metaweb (appendix figures one and two).
I've also been exploring the data available from Julian Resasco. The
primary constraint on further progress is that we are waiting on the
finalization of a data sharing agreement with RMBL.

\hypertarget{chapter-three-optimizing-corridor-placement-against-ecological-dynamics}{%
\section{Chapter Three: Optimizing corridor placement against ecological
dynamics}\label{chapter-three-optimizing-corridor-placement-against-ecological-dynamics}}

\hypertarget{objective-2}{%
\subsection{Objective}\label{objective-2}}

As land-use change has caused many habitats to become fragmented and
patchy, promoting landscape connectivity has become of significant
interest to mitigate the effects of this change on Earth's biodiversity.
However, the practical realities of conservation mean that there is a
limitation on how much we can modify landscapes in order to do this. So
what is the best place to put a corridor given a constraint on how much
surface-area you can change in a landscape? This is the question this
chapter seeks to answer. Models for inferring corridor locations have
been developed, but are limited in that are not developed around
promoting some element of ecosystem function, but instead by trying to
find the path of least resistance in an existing landscape from a
derived resistance surface (e.g. Peterman 2018). This chapter proposes a
general algorithm for optimizing spatial restoration effort to move a
measurement of ecosystem functioning toward a target state.

\hypertarget{methods-2}{%
\subsection{Methods}\label{methods-2}}

\begin{figure}
\hypertarget{fig:ch3}{%
\centering
\includegraphics{./figures/ch3.png}
\caption{A conceptual example of how we go from a map of landcover with
a set of points where occurrence of a species has been recorded (top
left), to the set of all possible landscape modifications (green box,
where each point in the green box is a unique landscape modification,
with three examples shown as pink, red and orange dots), to computing
resistance surfaces based on proposed landscape modifications (center)
and then simulate the distribution of extinction times for a
metapopulation in this new landscape (right).}\label{fig:ch3}
}
\end{figure}

We propose various landscape modifications that alter the cover of a
landscape, represented as a raster. We then compute a new resistance
surface based on the proposed landscape modification using Circuitscape
(McRae \emph{et al.} 2008), and based on the values of resistance to
dispersal between pair of locations we simulate spatially-explicit
metapopulation dynamics (Hanski \& Ovaskainen 2000; Ovaskainen \emph{et
al.} 2002) to estimate a distribution of time-until-extinction for each
landscape modification. The largest challenge in implementing this
algorithm is the space of potential modifications grows as \(O((nm)!)\)
for an \(n\) by \(m\) raster. For most actual landscapes to which we
wish to apply this method, the set of possible modifications becomes
uncomputably large, so we use simulated annealing to explore the search
space of possible modifications to estimate the modification that
maximizes the time-until extinction of simulated metapopulation dynamics
under that hypothetical modified landscape.

The biggest challenge in implementing simulated annealing in this
context is defining a proposal function for landscape modifications. At
the moment this is done by computing the minimum-spanning-tree (MST) of
the spatial nodes (locations where occurrence has been observed), and
then proposing corridors that only connect nodes that are already
connected in the MST. The primary reason for doing this is the cut down
the size of the solution space to enable quicker convergence, although
the final software that implements this algorithm will enable
alternative methods of proposing modifications.

The goal output of this chapter is not only to provide a set of discrete
corridor options, but also to rank the cells in the raster by priority
based on how many times they are included in the distribution of
``good'' corridors after simulated annealing has converged. Further, the
final component of this chapter is measuring the effect of land-use
change on the robustness of the optimized corridor by simulating various
neutral models of urban and agricultural sprawl, and determining if the
proposed modifications still maximize time-until-extinction when the
landcover in the landscape is not static.

\hypertarget{progress-2}{%
\subsection{Progress}\label{progress-2}}

Currently I have an algorithm for proposing landscape modifications and
a simple implementation of simulated annealing. The only gap left is
implementing Circuitscape estimation of resistance surfaces and running
benchmarking tests for the resulting chapter.

\hypertarget{chapter-four-metacommunitydynamics.jl-a-virtual-laboratory-for-community-ecology}{%
\section{Chapter Four: MetacommunityDynamics.jl: a virtual laboratory
for community
ecology}\label{chapter-four-metacommunitydynamics.jl-a-virtual-laboratory-for-community-ecology}}

\hypertarget{objective-3}{%
\subsection{Objective}\label{objective-3}}

The final chapter consists of a collection of modules in the
\texttt{Julia} language for different aspects of community ecology,
including most of the code used for the preceding chapters. Indeed
\texttt{MetacommunityDynamics.jl} (MCD.jl) is at the center of this set
of tools, but due to the nature of the Julia language, MCD.jl is
inter-operable with several existing packages within the
\texttt{EcoJulia} organization, including several to which I have
contributed. We need a software library like this to generate synthetic
data from a \emph{known} set of mechanisms and parameters to test our
methods for parameter inference and forecasting on this \emph{known}
system to assess the effectiveness of these inference and forecasting
methods.

\begin{figure}
\hypertarget{fig:software}{%
\centering
\includegraphics{./figures/ch4.png}
\caption{The structure of the software libraries used as part of
MCD.jl}\label{fig:software}
}
\end{figure}

\hypertarget{methods-3}{%
\subsection{Methods}\label{methods-3}}

A diagram showing the relation between these packages is shown in
fig.~\ref{fig:software}. \texttt{MetacommunityDynamics.jl} is built on
\texttt{DynamicGrids.jl}, a library for high-performance gridded
simulations in the \texttt{Julia} language, and Dispersal.jl (Maino
\emph{et al.} 2021), and extension of \texttt{DynamicGrids.jl}
specifically for modeling organism dispersal. It also contains
integrations with \texttt{EcologicalNetworks.jl} (Poisot \emph{et al.}
2019) to generate metawebs, or can use empirical networks from Mangal.jl
(Banville \emph{et al.} 2021). It implements the general framework for
community dynamics proposed by Vellend (2010), where all community
processes can divided into four categories: selection, dispersal, drift,
and speciation.

\hypertarget{results-2}{%
\subsection{Results}\label{results-2}}

In fig.~\ref{fig:foodwebtraj} we see a sample output of simulated
food-web dynamics for a metaweb of 100 species generated using the
minimum-potential-niche model (Allesina \emph{et al.} 2008) with
connectance \(C=0.05\) and forbidden-link probability of \(0.5\). The
dynamics change according to a Lotka-Volterra functional response,
dispersal with distance inversely proportional to trophic-level, linear
mortality, and logistic growth for any species at the producer
trophic-level.

\begin{figure}
\hypertarget{fig:foodwebtraj}{%
\centering
\includegraphics{./figures/foodwebtraj.png}
\caption{Sample output of simulated food web dynamics from
MetacommunityDynamics.jl. Timestep (x-axis), and biomass of each species
(y-axis).}\label{fig:foodwebtraj}
}
\end{figure}

\hypertarget{progress-3}{%
\subsection{Progress}\label{progress-3}}

The software as it exists is capable of simulating the biomass dynamics
of arbitrarily large food-webs using Lotka-Volterra, Holling Type-II, or
Holling Type-III functional responses. It currently has methods to
implement Gaussian drift, and various forms of dispersal via
\texttt{Dispersal.jl}. Also functional are occupancy dynamics for Levins
metapopulations (Levins 1969), and spatially explicit Hanski-Ovaskainen
metapopulations (Hanski \& Ovaskainen 2000; Ovaskainen \emph{et al.}
2002). This is most of what needs to exist for the preceding chapters.
In-progress functionality includes selection (which affects growth-rate)
on arbitrary environmental variables in progress, as well as traits.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Developing a system for global biodiversity monitoring is an imperative
to mitigate biodiversity loss and its impacts on humanity. In my thesis
I hope to provide a template for the digital infrastructure that enables
the pipeline from data collection, to forecast, to actionable
information, both through software that can be used to solve these
problems (chapters one, three, four), and vignettes of how these
software can be applied (chapters one, two).

Biodiversity science would be wise to use the success of
numerical-weather prediction and the Earth monitoring system that
supports it as a template. Further, we should embrace methodological
advances in computational statistics that can enable more robust
prediction of the dynamics of complex systems than is possible using the
tools of analytic models---not just because they can provide more
efficacious predictions, but also because they enable a more direct
representation of uncertainty, which is crucial for developing
ecological forecasting models which provide actionable information to
stakeholders, and to find a way for humans to live sustainably on Earth,
the only planet on which life has ever been known to occur.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Allesina2008GenMod}{}%
Allesina, S., Alonso, D. \& Pascual, M. (2008). A General Model for Food
Web Structure. \emph{Science}, 320, 658--661.

\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
2721.

\leavevmode\hypertarget{ref-Bascompte2007PlaMut}{}%
Bascompte, J. \& Jordano, P. (2007). Plant-Animal Mutualistic Networks:
The Architecture of Biodiversity. \emph{Annual Review of Ecology,
Evolution, and Systematics}, 38, 567--593.

\leavevmode\hypertarget{ref-Bauer2015QuiRev}{}%
Bauer, P., Thorpe, A. \& Brunet, G. (2015). The quiet revolution of
numerical weather prediction. \emph{Nature}, 525, 47--56.

\leavevmode\hypertarget{ref-Beckage2011LimPre}{}%
Beckage, B., Gross, L.J. \& Kauffman, S. (2011). The limits to
prediction in ecological systems. \emph{Ecosphere}, 2, art125.

\leavevmode\hypertarget{ref-CaraDonna2017IntRew}{}%
CaraDonna, P.J., Petry, W.K., Brennan, R.M., Cunningham, J.L.,
Bronstein, J.L., Waser, N.M., \emph{et al.} (2017). Interaction rewiring
and the rapid turnover of plantpollinator networks. \emph{Ecology
Letters}, 20, 385--394.

\leavevmode\hypertarget{ref-Chen2019RevCom}{}%
Chen, Y., Angulo, M.T. \& Liu, Y.-Y. (2019). Revealing Complex
Ecological Dynamics via Symbolic Regression. \emph{BioEssays}, 41,
1900069.

\leavevmode\hypertarget{ref-Dietze2017PreEco}{}%
Dietze, M.C. (2017). Prediction in ecology: A first-principles
framework. \emph{Ecological Applications}, 27, 2048--2060.

\leavevmode\hypertarget{ref-Gravel2019BriElt}{}%
Gravel, D., Baiser, B., Dunne, J.A., Kopelke, J.-P., Martinez, N.D.,
Nyman, T., \emph{et al.} (2019). Bringing Elton and Grinnell together: A
quantitative framework to represent the biogeography of ecological
interaction networks. \emph{Ecography}, 42, 401--415.

\leavevmode\hypertarget{ref-Hadfield2014TalTwo}{}%
Hadfield, J.D., Krasnov, B.R., Poulin, R. \& Nakagawa, S. (2014). A Tale
of Two Phylogenies: Comparative Analyses of Ecological Interactions.
\emph{The American Naturalist}, 183, 174--187.

\leavevmode\hypertarget{ref-Hanski2000MetCap}{}%
Hanski, I. \& Ovaskainen, O. (2000). The metapopulation capacity of a
fragmented landscape. \emph{Nature}, 404, 755--758.

\leavevmode\hypertarget{ref-Hill2004ArcEar}{}%
Hill, C., DeLuca, C., Balaji, Suarez, M. \& Da Silva, A. (2004). The
architecture of the Earth System Modeling Framework. \emph{Computing in
Science Engineering}, 6, 18--28.

\leavevmode\hypertarget{ref-Hubbell2001UniNeu}{}%
Hubbell, S.P. (2001). \emph{The unified neutral theory of biodiversity
and biogeography}. Monographs in population biology. Princeton
University Press, Princeton.

\leavevmode\hypertarget{ref-Karger2017CliHig}{}%
Karger, D.N., Conrad, O., Böhner, J., Kawohl, T., Kreft, H., Soria-Auza,
R.W., \emph{et al.} (2017). Climatologies at high resolution for the
earth's land surface areas. \emph{Scientific Data}, 4, 170122.

\leavevmode\hypertarget{ref-Levin1992ProPat}{}%
Levin, S.A. (1992). The Problem of Pattern and Scale in Ecology: The
Robert H. MacArthur Award Lecture. \emph{Ecology}, 73, 1943--1967.

\leavevmode\hypertarget{ref-Levins1969DemGen}{}%
Levins, R. (1969). Some Demographic and Genetic Consequences of
Environmental Heterogeneity for Biological Control. \emph{Bulletin of
the Entomological Society of America}, 15, 237--240.

\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
MacDonald, A.A.M., Banville, F. \& Poisot, T. (2020). Revisiting the
Links-Species Scaling Relationship in Food Webs. \emph{Patterns}, 1.

\leavevmode\hypertarget{ref-Maino2021PreGlo}{}%
Maino, J.L., Schouten, R. \& Umina, P. (2021). Predicting the global
invasion of Drosophila suzukii to improve Australian biosecurity
preparedness. \emph{Journal of Applied Ecology}, 58, 789--800.

\leavevmode\hypertarget{ref-Makiola2020KeyQue}{}%
Makiola, A., Compson, Z.G., Baird, D.J., Barnes, M.A., Boerlijst, S.P.,
Bouchez, A., \emph{et al.} (2020). Key Questions for Next-Generation
Biomonitoring. \emph{Frontiers in Environmental Science}, 7.

\leavevmode\hypertarget{ref-McRae2008UsiCir}{}%
McRae, B.H., Dickson, B.G., Keitt, T.H. \& Shah, V.B. (2008). Using
Circuit Theory to Model Connectivity in Ecology, Evolution, and
Conservation. \emph{Ecology}, 89, 2712--2724.

\leavevmode\hypertarget{ref-Ovaskainen2002MetMod}{}%
Ovaskainen, O., Sato, K., Bascompte, J. \& Hanski, I. (2002).
Metapopulation Models for Extinction Threshold in Spatially Correlated
Landscapes. \emph{Journal of Theoretical Biology}, 215, 95--108.

\leavevmode\hypertarget{ref-Ovaskainen2002MetMod}{}%
Ovaskainen, O., Sato, K., Bascompte, J. \& Hanski, I. (2002).
Metapopulation Models for Extinction Threshold in Spatially Correlated
Landscapes. \emph{Journal of Theoretical Biology}, 215, 95--108.

\leavevmode\hypertarget{ref-Pennekamp2019IntPre}{}%
Pennekamp, F., Iles, A.C., Garland, J., Brennan, G., Brose, U., Gaedke,
U., \emph{et al.} (2019). The intrinsic predictability of ecological
time series and its potential to guide forecasting. \emph{Ecological
Monographs}, 89, e01359.

\leavevmode\hypertarget{ref-Petchey2015EcoFor}{}%
Petchey, O.L., Pontarp, M., Massie, T.M., Kéfi, S., Ozgul, A.,
Weilenmann, M., \emph{et al.} (2015). The ecological forecast horizon,
and examples of its uses and determinants. \emph{Ecology Letters}, 18,
597--611.

\leavevmode\hypertarget{ref-Peterman2018ResRP}{}%
Peterman, W.E. (2018). ResistanceGA: An R package for the optimization
of resistance surfaces using genetic algorithms. \emph{Methods in
Ecology and Evolution}, 9, 1638--1647.

\leavevmode\hypertarget{ref-Poisot2019EcoJl}{}%
Poisot, T., Bélisle, Z., Hoebeke, L., Stock, M. \& Szefer, P. (2019).
EcologicalNetworks.jl: Analysing ecological networks of species
interactions. \emph{Ecography}, 42, 1850--1861.

\leavevmode\hypertarget{ref-Strydom2021FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M.J., \emph{et al.} (2021a). Food web reconstruction through
phylogenetic transfer of low-rank network representation.

\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
Strydom, T., Catchen, M.D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \emph{et al.} (2021b). \emph{A Roadmap Toward
Predicting Species Interaction Networks (Across Space and Time)}
(Preprint). EcoEvoRxiv.

\leavevmode\hypertarget{ref-Thompson2000ResSol}{}%
Thompson, R.M. \& Townsend, C.R. (2000). Is resolution the solution?:
The effect of taxonomic resolution on the calculated properties of three
stream food webs. \emph{Freshwater Biology}, 44, 413--422.

\leavevmode\hypertarget{ref-Urban2021CodLif}{}%
Urban, M.C., Travis, J.M.J., Zurell, D., Thompson, P.L., Synes, N.W.,
Scarpa, A., \emph{et al.} (2021). Coding for Life: Designing a Platform
for Projecting and Protecting Global Biodiversity. \emph{BioScience}.

\leavevmode\hypertarget{ref-Vellend2010ConSyn}{}%
Vellend, M. (2010). Conceptual Synthesis in Community Ecology. \emph{The
Quarterly Review of Biology}, 85, 183--206.

\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
Williams, R.J. \& Martinez, N.D. (2000). Simple rules yield complex food
webs. \emph{Nature}, 404, 180--183.

\end{CSLReferences}

\end{document}